\documentclass[a4paper,12pt]{article}
\usepackage{fullpage}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{authblk}
\usepackage{todonotes}
\usepackage{caption}
\usepackage{url}
\usepackage{float}
\usepackage{sectsty}
\usepackage{pdfpages}
\usepackage[section]{placeins}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}

\usepackage{setspace}
\usepackage[toc,page]{appendix}
\usepackage{framed}
\usepackage{geometry}

\usepackage{alltt}
\usepackage{subfig}

% Change section fonts
\allsectionsfont{\sffamily}

% For code box
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{%
  \parbox{\textwidth}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}\vskip-4pt}}
  \captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
  \lstset{frame=lrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}
% End code box

\usepackage{cite}

% General parameters, for ALL pages:
\renewcommand{\topfraction}{0.9}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}	% max fraction of floats at bottom
% Parameters for TEXT pages (not float pages):
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4} % 2 may work better
\setcounter{dbltopnumber}{2} % for 2-column pages

\addtolength{\topmargin}{0.5in}

\usepackage{fancyvrb}

\usepackage{tikz} \usetikzlibrary{trees}
\usepackage{hyperref} % should always be the last package

% useful colours (use sparingly!):
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\green}[1]{{\color{green}#1}}
\newcommand{\red}[1]{{\color{red}#1}}

% useful wrappers for algorithmic/Python notation:
\newcommand{\length}[1]{\text{len}(#1)}
\newcommand{\twodots}{\mathinner{\ldotp\ldotp}} % taken from clrscode3e.sty
\newcommand{\Oh}[1]{\mathcal{O}\left(#1\right)}

% useful (wrappers for) math symbols:
\newcommand{\Cardinality}[1]{\left\lvert#1\right\rvert}
\newcommand{\Ceiling}[1]{\left\lceil#1\right\rceil}
\newcommand{\Floor}[1]{\left\lfloor#1\right\rfloor}
\newcommand{\Iff}{\Leftrightarrow}
\newcommand{\Implies}{\Rightarrow}
\newcommand{\Intersect}{\cap}
\newcommand{\Sequence}[1]{\left[#1\right]}
\newcommand{\Set}[1]{\left\{#1\right\}}
\newcommand{\SetComp}[2]{\Set{#1\SuchThat#2}}
\newcommand{\SuchThat}{\mid}
\newcommand{\Tuple}[1]{\langle#1\rangle}
\newcommand{\Union}{\cup}
\usetikzlibrary{positioning,shapes,shadows,arrows}

\renewcommand{\abstractname}{Elfin - URL Shortener}

\title{\textbf{Elfin - URL Shortener}}
\author{Lukas Klingsbo}
\begin{document}
\maketitle

\begin{abstract}
I was given this task by Uprise, as an exercise to show my capabilities.\\
This URL shortener is built with Scala 2.11.7 and the Lift 2.6 Framework.
The result can be seen on \url{http://elfin.se}

\textbf{Disclaimer:} This is the first Lift project that I have 
coded so please forgive any horrible breaking of convention 
that I might have unintentionally made. 
\begin{center}
If you have any questions,\\
feel free to contact me on:\\
\vspace{10pt}
Lukas Klingsbo\\
+46737-42 43 45\\
lukas.klingsbo@gmail.com\\
\end{center}

\section{Background}
The task was to make a scalable URL shortener that deterministically shortened 
the URL's fed to it.

\section{Design decisions}
The shortener core code was written with the help of MurmurHash3 which I use 
to convert the $URL \Rightarrow Int \Rightarrow String \Rightarrow URL$. 

Another design choice that was considered was using an auto incremented value 
in a database that was to be converted to a string. This would have yielded 
shorter strings and avoided the problem with collisions that occur when using 
short hashes. The first design choice was chosen as the exercise stated that 
the same link preferably always should produce the same short URL. It also 
gives the advantage that the keys does not have to be synchronized on creation 
to the distributed KV-store, they still have to be synchronized for fetching 
though.

\subsection{Database}
As the task was to make a scalable URL shortener with high availability, the 
choice to use a distributed database (in production) was made.
Two Key-Value stores were looked into, Riak and Redis. Redis was choosen simply 
because it had better support for Scala. When deploying Redis for production 
for this project I recommend using a single master + replicas and then use 
Sentinel for automatic failover. Which will make the system scalable and 
extremely fault tolerant as there is no sharding, all data will exist on all 
of the nodes and when the master goes down a new one is selected to become 
master by Sentinel.

As the algorithm for determining short URL's is deterministic as many web 
servers as needed can be used individually with a load balancer to scale 
the application.

\subsection{Variable Key length}
When using such short hashes that are used here, then collisions are 
unavoidable (See the pigeonhole principle). So consider how much entropy 
you need for the intended usage and set a fixed length of the key before 
deploying this application.

\section{Setup}
To try this application you can either use SBT directly or package a WAR 
file and deploy it on a web server and java servlet containers like Jetty 
or Tomcat. An instance of Redis is also needed to properly run the 
application.

\section{Todo}
In a future version the history of shortened URL's should be provided to 
the user by linking the URL's to the users session key. 


\pagenumbering{gobble} % Remove page numbers (and reset to 1)
\end{abstract}
\end{document}
